---
title: "STAT7123/STAT8123"
author: "Umut Demirhan"
date: "Due 11:55 pm, Friday September 29th, 2023"
output:
  pdf_document:
    keep_tex: no
    includes: null
  word_document: default
subtitle: Statistical Graphics Assignment 2
editor_options:
  markdown:
    wrap: 72
---

\section{Question 1}

a)  The total number of offences by local government area (LGA).

```{r, echo=FALSE, include=FALSE, results='hide'}
# Load the necessary libraries
library(tidyverse)
library(ggplot2)
library(dplyr)
library(sf)
library(purrr)
library(scales)
library(stringr)
library(tidyr)
library(knitr)
library(kableExtra)
```

```{r, echo=FALSE}
# Read  the CSV data
data <- read.csv("penalty_notice.csv")
# Calculate the total number of offences by LGA 
data_summarized <- data %>%
  group_by(Offence_LGA) %>%
  summarise(total_offences = n()) 

#print(data_summarized)
#count(data_summarized)
head(data_summarized)
```

Given the summary provided, one challenging feature regarding plotting
this data is the significant disparity in the total number of offences
between different Local Government Areas (LGA). Some LGAs have over a
hundred offences, while many others have only one, leading to a high
variance in the data which can make visualization unclear or misleading
if not handled properly.

Secondly, there are a considerable number of LGAs listed, which could
result in a cluttered and unreadable plot, especially if every LGA is
labeled, hence requiring careful consideration of labeling, positioning,
and choice of plot type.

Lastly, since this is count data, choosing the right type of plot is
crucial. If a choropleth map is not the choice, bar plot might be the
most straightforward choice, but with the disparity in the number of
offences, using a logarithmic scale might be necessary to visualize low
and high count LGAs together effectively.

b)  Which LGA has the most offences?

```{r, echo=FALSE,message=FALSE, warning=FALSE}
# This code sorts the summarized data to find the LGA with the most offences.

# Sorting to find the LGA with the most offences
data_sorted <- data_summarized %>% arrange(desc(total_offences))
top_LGA <- data_sorted[1, c("Offence_LGA", "total_offences")]

# Displaying the LGA with the most offences along with the number of offences
top_LGA <- data.frame(top_LGA)
#top_LGA
```

The LGA with the most offences is stored in the variable top_LGA.
According to the sorted data, the LGA with the most offences is
Canterbury-Bankstown, having 349 total offences. \newpage c) A bar chart
to display a subset of the LGAs

```{r, echo=FALSE,message=FALSE, warning=FALSE}
# Filter to keep only the top 20 LGAs 
top_LGAs <- data_sorted %>% 
  filter(row_number() <= 20)

# Create the plot 
ggplot(top_LGAs, aes(x = reorder(Offence_LGA, total_offences), 
                     y = total_offences)) +
  geom_bar(stat = "identity", fill = "steelblue", alpha = 0.5) +
  geom_text(aes(label = total_offences), hjust = 1.5, color = "black",
            size = 2.4) + 
  coord_flip() +
  xlab("Local Government Area") +
  ylab("Total Number of Offences") +
  ggtitle("Top 20 LGAs by Number of Offences") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 8))
```

The LGA with the most offences is Canterbury-Bankstown, with a total of
349 offences. This is significantly higher than the second and
third-ranking LGAs, Burwood and Blacktown, which have 127 and 107
offences respectively. The chart also shows that the number of offences
varies considerably among these top 20 LGAs, ranging from a high of 349
in Canterbury-Bankstown to a low of 13 in Inner West.

The horizontal bars are color-coded in steel blue and are ordered in
descending fashion, making it easy to identify LGAs with more frequent
offences. Text labels indicate the exact number of offences for each
LGA, providing a quick numerical reference for the viewer. Overall, the
chart provides a succinct yet comprehensive overview of food safety
offences in the top 20 LGAs, highlighting areas that may require more
stringent regulatory attention. \newpage

\section{Question 2}

a)  A plot for number of offences by time

```{r, echo=FALSE,message=FALSE, warning=FALSE}
# Loading the package "zoo"
library(zoo)

# Converting year and month into a unified time variable
data$time <- as.yearmon(paste(data$Year, data$Month), "%Y %m")

# Calculating the number of offences by month per year
data_by_time <- data %>%
  group_by(time) %>%
  summarise(total_offences = n())

# Creating the plot for number of offences by time
ggplot(data_by_time, aes(x = as.Date(time), y = total_offences)) +  
  # Convert time to Date for ggplot
  geom_line(color = "blue") +
  geom_point(color = "red") +
  xlab("Time (Year and Month)") +
  ylab("Number of Offences") +
  ggtitle("Number of Offences Over Time") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
    panel.grid.major.x = element_line(color = "grey90", size = 0.3),
    panel.grid.minor.x = element_blank(),
    plot.title = element_text(hjust = 0.5),  
    plot.title.position = "plot"  
     ) +
    scale_x_date(breaks = as.Date(data_by_time$time), date_labels = "%Y-%m")  

```

b)  Interpretation of the plot

The line chart reveals a dramatic fluctuation in the number of offences
from December 2021 to June 2023. Notably, there are peaks observed
around May and June of 2022 and troughs in early 2022 and June 2023.
While there is no pattern for seasonality, the drastic increase in May
2022 and decline in offences in recent months suggests that other
factors may also be at play such as lifted quarantine in February 2022
in Sydney. Overall, the chart illustrates that the number of offences
has not remained consistent over time and warrants further investigation
to understand these variations.

c)  A plot for average amount payable by time

```{r, echo=FALSE,message=FALSE, warning=FALSE}
# Calculate the average amount payable by date
data_avg_amount <- data %>%
  group_by(time) %>%
  summarise(average_amount = mean(Amount_Payable))

# Create the plot for average amount payable by time
ggplot(data_avg_amount, aes(x = as.Date(time), y = average_amount)) +
  geom_line(color = "#5F9EA0") +
  geom_point(color = "#00008B") +
  xlab("Time (Year and Month)") +
  ylab("Average Amount Payable ($)") +
  ggtitle("Average Amount Payable Over Time") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
    panel.grid.major.x = element_line(color = "grey90", size = 0.3),
    panel.grid.minor.x = element_blank(),
    plot.title = element_text(hjust = 0.5),
    plot.title.position = "plot"  
  ) +
  scale_x_date(breaks = as.Date(data_by_time$time), date_labels = "%Y-%m")
```

d)  Interpretation of the plot

The plot of average amounts payable over time predominantly exhibits no
discernible seasonal or linear trends, with the exception of a short
period from December 2021 to March 2022, which may reflect increased
monitoring coinciding with a spike in offences. The data points
fluctuate up and down, but these movements are somewhat sporadic and do
not follow a consistent pattern across the observed period. While there
are some high and low points, they don't occur with the regularity that
would suggest seasonality. Therefore, the plot reveals variability but
does not provide evidence for a distinct trend or seasonality in the
average amounts payable for offences. \newpage

\section{Question 3}

a)  A violin plot for penalty amounts by Issuing Authority

```{r, echo=FALSE,message=FALSE, warning=FALSE}
# Include the package to plot with a colorblind-friendly palette
library(viridis)
library(RColorBrewer)
# Recoding the 'Issuing_Authority' column to separate 
# 'NSW Food Authority' from all other authorities categorized as 'Councils'
data$Revised_Issuing_Authority <- 
  ifelse(data$Issuing_Authority == 
           "NSW Food Authority", "NSW Food Authority", "Councils")

# Create a violin plot
ggplot(data, aes(x = Revised_Issuing_Authority, y = Amount_Payable)) +
  geom_violin(aes(fill = Revised_Issuing_Authority), trim = FALSE, alpha = 0.2)+ 
  geom_jitter(aes(color = Revised_Issuing_Authority), width = 0.1, size = 0.5) + 
  xlab("Issuing Authority") +
  ylab("Amount Payable ($)") +
  ggtitle("Violin Plot of Penalty Amounts by Issuing Authority") +
  theme_minimal() +
  theme(
    
    plot.title = element_text(hjust = 0.5),
    plot.title.position = "plot",
    legend.position = "bottom",  
    legend.title = element_blank()  
  ) +
  scale_fill_viridis_d(option = "C", end = 0.9, begin = 0.1, name = NULL) +  
  scale_color_viridis_d(option = "C", end = 0.9, begin = 0.1, name = NULL)  
# ChatGPT helped to manage this properties of "viridis"
```

b)  Interpretation of the plot

-   There are seven distinct values for amount payable, and the violin
    plot reveals those penalty amounts for both the councils and the NSW
    Food Authority.
-   Councils frequently issue penalties below \$500 (\$440 and \$330)
    and mostly around \$800 (\$880, \$770, and \$660), whereas the NSW
    Food Authority focuses on the \$800 range while alsovbeing more
    diverse in penalties.
-   The NSW Food Authority's plot is more balanced, indicating a greater
    diversity in penalty amounts, including those over \$1000.
-   The general pattern for both indicates that penalties over \$1000
    (\$1540 and \$1320) and \$660 are sparse, penalties below \$500
    (\$440 and \$330) are pretty popular, and penalties around \$800
    (\$880, \$770) are the most frequent.

```{=tex}
\newpage
\section{Question 4}
```
a)  A choropleth map that shows the number of penalties by LGA in NSW

```{r, echo=FALSE,message=FALSE, warning=FALSE, results='hide'}
# Read the shapefile
shapefile_data <- st_read("LGA_2021_AUST_GDA2020_SHP/LGA_2021_AUST_GDA2020.shp")
# Filter for New South Wales
shapefile_data_nsw <- shapefile_data %>% filter(STE_NAME21 == "New South Wales")
```

```{r, echo=FALSE,message=FALSE, warning=FALSE,height=5, width=6}
# Summarizing data by LGA
data_summarized <- data %>% 
  group_by(Offence_LGA) %>% 
  summarise(total_offences = n()) %>% 
  rename(NAME = Offence_LGA)

# Due to Data mismatching for some LGAs, conducting following operation.

# Making LGA names compatible for merging
shapefile_data_nsw$LGA_NAME21 <- tolower(trimws(shapefile_data_nsw$LGA_NAME21))
data_summarized$NAME <- tolower(trimws(data_summarized$NAME))

# Removing "(nsw)" from the end of the LGA names in shapefile_data_nsw 
shapefile_data_nsw$LGA_NAME21 <- sub("\\(nsw\\)$", "", 
                                     shapefile_data_nsw$LGA_NAME21)

# Removing "city of " from the beginning of the LGA names in data_summarized
data_summarized$NAME <- sub("^city of ", "", data_summarized$NAME)

# Trim extra whitespace for merging
data_summarized$NAME <- trimws(data_summarized$NAME)
shapefile_data_nsw$LGA_NAME21 <- trimws(shapefile_data_nsw$LGA_NAME21)

# Manually adjusting the special cases in shapefile_data_nsw
shapefile_data_nsw$LGA_NAME21[shapefile_data_nsw$LGA_NAME21 == 
                      "queanbeyan-palerang regional"] <-"queanbeyan-palerang"
shapefile_data_nsw$LGA_NAME21[shapefile_data_nsw$LGA_NAME21 == 
                                "sutherland shire"] <- "sutherland"
shapefile_data_nsw$LGA_NAME21[shapefile_data_nsw$LGA_NAME21 == 
                                "tamworth regional"] <- "tamworth"
shapefile_data_nsw$LGA_NAME21[shapefile_data_nsw$LGA_NAME21 ==
                                "the hills shire"] <- "the hills"

# Merging spatial data with summarized data
merged_data <- left_join(shapefile_data_nsw, data_summarized,
                         by = c("LGA_NAME21" = "NAME"))
# Used ChatGPT for some of the above operation(debugging etc.)

# Running the unmatched checks
#unmatched_shapefile_LGAs <- setdiff(shapefile_data_nsw$LGA_NAME21,
#                                    merged_data$LGA_NAME21)
#print("Unmatched LGAs in shapefile_data_nsw:")
#print(unmatched_shapefile_LGAs)

#unmatched_summary_LGAs <- setdiff(data_summarized$NAME, merged_data$LGA_NAME21)
#print("Unmatched LGAs in data_summarized:")
#print(unmatched_summary_LGAs)

# Creating the choropleth map
ggplot(data = merged_data, mapping = aes(fill = total_offences)) +
  geom_sf(color = "white", size = 0.1) + 
  scale_fill_gradient(
    low = "lightblue", 
    high = "darkblue", 
    na.value = "grey70",
    name = "Total Offences",
    breaks = pretty_breaks(n = 5), 
    labels = scales::comma
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
    axis.title.x = element_text(size = 10, face = "bold"),
    axis.title.y = element_text(size = 10, face = "bold"),
    legend.position = "bottom",
    legend.title = element_text(size = 10, face = "bold"),
    legend.text = element_text(size = 8),
    legend.key.width = unit(1.5, "cm")
  ) +
  labs(
    title = "Number of Penalties by LGA in NSW",
    x = "Longitude",
    y = "Latitude"
  ) +
  coord_sf(xlim = c(140, 155),ylim = c(-38,-28))
```
\newpage
b)  A choropleth map focusing on Sydney region

```{r, echo=FALSE,message=FALSE, warning=FALSE,height=5, width=5}
# Creating the choropleth map focusing on Sydney region
ggplot(data = merged_data, mapping = aes(fill = total_offences)) +
  geom_sf(color = "white", size = 0.1) + 
  scale_fill_gradient(
    low = "lightblue", 
    high = "darkblue", 
    na.value = "grey70",
    name = "Total Offences",
    breaks = pretty_breaks(n = 5), 
    labels = scales::comma
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
    axis.title.x = element_text(size = 10, face = "bold"),
    axis.title.y = element_text(size = 10, face = "bold"),
    legend.position = "bottom",
    legend.title = element_text(size = 10, face = "bold"),
    legend.text = element_text(size = 8),
    legend.key.width = unit(1.5, "cm")
  ) +
  labs(
    title = "Number of Penalties by LGA in Sydney Region",
    x = "Longitude",
    y = "Latitude"
  ) +
  coord_sf(xlim = c(150.5, 151.5), ylim = c(-34.2, -33.5))  
# Restricting longitude and latitude for Sydney region

```

c)  Why focus on the Sydney region for number of penalty per LGA?

Focusing on the Sydney region allows us to highlight densely populated
areas and, therefore, more meaningful variations in penalty numbers, as
it contains the areas with the highest number of penalties, whereas
other regions in the expansive NSW territory often have so few penalties
that they only appear in the lightest colours on a statewide map.

d)  Summary of the main results of the choropleth plot presented.

Within the Sydney-centric map, Canterbury-Bankstown emerges prominently
due to its darker hue, signifying the highest number of offences among
the Local Government Areas (LGAs). Although some other regions like
Burwood and Blacktown register elevated offence levels, they are
markedly lower than those in Canterbury-Bankstown. Intriguingly, despite
hosting the highest number of restaurants, the City of Sydney doesn't
reflect a proportional penalty rate, emphasizing the urgent need for
further investigations of the reason behind it and/or targeted
interventions in areas like Canterbury-Bankstown.

```{=tex}
\newpage
\section{Question 5}
```
a)  Data Simplification Steps for Nature_of_Offence_Full for Graphical
    Presentation

**Objective:**

To refine the lengthy descriptions in `Nature_of_Offence_Full` to a
simplified, categorized format suitable for graphical representation,
without losing the crucial information.

**Proposed Steps:**

-   **Standardization:**

    -   Ensure uniformity in text to prevent redundancy and
        discrepancies during categorization.
    -   Convert the text in `Nature_of_Offence_Full` to lowercase for
        uniformity.
    -   Standardize variants of similar phrases to a common
        representation, such as different forms of 'previous warning
        given'.

-   **Categorization and Simplification:**

    -   Generalize and simplify the nature of offences to manageable categories
        for effective graphical representation.

    -   Identify key terms or phrases from the detailed descriptions to create
        simplified categories, e.g., "Illegal Preservative", "Cleanliness of 
        Utensils".

    -   Assign each entry a simplified category based on the identified key
        terms or phrases. Classify unmatching entries as “Other”.


-   **Creating Indicator Variables:**

    -   Extract additional meaningful information from the detailed
        descriptions.
    -   Create an indicator variable to denote whether previous warnings
        were mentioned,providing insights into the compliance history of
        the entity involved.

-   **Data Aggregation:**

    -   Facilitate the creation of graphical representations by
        summarizing the data.
    -   Aggregate the data based on the new simplified categories and
        indicator variables to find the frequency of each type of
        offence.

\newpage

b)  Data Simplification process for Nature_of_Offence_Full for Graphical
    Presentation and Three Statistical Graphics

```{r, echo=FALSE,message=FALSE, warning=FALSE}
# Define the pattern that matches all variants of 'previous warning given'
data <- data %>%
  mutate(Nature_of_Offence_Full = gsub("previous waning given|previous warnings 
                      given|previous waning given", "previous warning given",
                      Nature_of_Offence_Full, ignore.case = TRUE))

# Converting Nature_of_Offence_Full to lower case for uniformity
data$Nature_of_Offence_Full <- tolower(data$Nature_of_Offence_Full)

# Create a new column for simplified categories
data$Simplified_Nature_of_Offence <- 
  ifelse(grepl("illegal preservative", data$Nature_of_Offence_Full),
         "Illegal Preservative",
ifelse(grepl("utensils", data$Nature_of_Offence_Full),
       "Cleanliness of Utensils",
ifelse(grepl("hand washing", data$Nature_of_Offence_Full),
       "Hand Washing Facilities",
ifelse(grepl("food safety supervisor", data$Nature_of_Offence_Full),
       "Food Safety Supervisor",
ifelse(grepl("food safety", data$Nature_of_Offence_Full),
       "Food Safety",
ifelse(grepl("food that is unsuitable", data$Nature_of_Offence_Full),
       "Unsuitable Food",
ifelse(grepl("unsafe", data$Nature_of_Offence_Full), "Unsafe Food",
ifelse(grepl("temperature", data$Nature_of_Offence_Full),
       "Temperature Control",
ifelse(grepl("equipment", data$Nature_of_Offence_Full),
       "Surface Contamination",
ifelse(grepl("pests", data$Nature_of_Offence_Full), "Pest Control",
ifelse(grepl("food contact surfaces", data$Nature_of_Offence_Full),
       "Food Contact Surfaces",
ifelse(grepl("food labelled in contravention", data$Nature_of_Offence_Full),
       "Labeling Issues",
ifelse(grepl("food handling", data$Nature_of_Offence_Full), "Handling Skills",
ifelse(grepl("store food", data$Nature_of_Offence_Full), "Food Storage",
ifelse(grepl("prohibition order", data$Nature_of_Offence_Full), "Legal Offense",
ifelse(grepl("licence", data$Nature_of_Offence_Full), "Licensing Issue", 
ifelse(grepl("design and construction", data$Nature_of_Offence_Full),
       "Design and Construction",
ifelse(grepl("contamina", data$Nature_of_Offence_Full),
       "Contamination Prevention",
ifelse(grepl("use-by date", data$Nature_of_Offence_Full), "Use-by Date",
ifelse(grepl("waste water disposal", data$Nature_of_Offence_Full), 
       "Waste Water Disposal",            
ifelse(grepl("adequate supply of water", data$Nature_of_Offence_Full),
       "Water Supply",         
ifelse(grepl("live animal", data$Nature_of_Offence_Full), 
       "Animals in Cooking Area",     
ifelse(grepl("garbage", data$Nature_of_Offence_Full), 
       "Garbage Management",     
ifelse(grepl("allergen advisory", data$Nature_of_Offence_Full), 
       "Allergen Advisory",                  
ifelse(grepl("designed", data$Nature_of_Offence_Full), "Design for Safety",                           ifelse(grepl("maintain the food premises", data$Nature_of_Offence_Full),
"Sanitation Failure",                        
ifelse(grepl("floors are able to be easily and effectively cleaned",
             data$Nature_of_Offence_Full), "Cleanliness of Floors",                                                      "Other")))))))))))))))))))))))))))
# Create a new column to indicate if previous warnings were given
data$Has_Warnings <- ifelse(grepl("previous warnings?|previous warning given", 
                                  data$Nature_of_Offence_Full), "Yes", "No")
```

Graphic 1: Bar Plot of Top 5 Most Common Offences per Year

```{r, echo=FALSE,message=FALSE, warning=FALSE,height=5, width=5}
# Count the frequency of each Simplified_Nature_of_Offence and Has_Warnings 
offence_count <- data %>%
  group_by(Simplified_Nature_of_Offence, Has_Warnings) %>%
  summarise(count = n()) %>%
  arrange(Simplified_Nature_of_Offence, desc(Has_Warnings))
# Generate the bar chart
ggplot(offence_count, aes(x = reorder(Simplified_Nature_of_Offence, -count), 
                          y = count, fill = as.factor(Has_Warnings))) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  coord_flip() +
  scale_fill_manual(values = c("#87CEFF", "#0000CD"), 
                    name = "Warnings Given Before", labels = c("Yes", "No")) +
  labs(title = "Frequency of Simplified Nature of Offence",
       subtitle = "Colored by warning status",
       x = "Offences",
       y = "Frequency") +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 12),
    plot.subtitle = element_text(size = 8,hjust = 0.5),
    legend.position = "right",
    legend.direction = "vertical",
    legend.text = element_text(size = 8),
    legend.title = element_text(size = 8),
    legend.key.size = unit(0.5, "cm")
  )
```

**Interpretation of the Frequency of Simplified Offences Bar Chart**

The bar chart provides a clear visualization of the frequency of various
offenses, distinguished by whether warnings were given before for the
same issue.

-   "Surface Contamination(Food Contact)" and "Sanitation Failure (Food
    Premises)" are predominant, with an almost equal distribution of
    warnings and non-warnings.

-   "Hand Washing Facilities," "Food Storage,", "Pest Control" and
    "Temperature Control" occur significantly, with variances in warning
    distributions.

-   Issues with "Illegal Preservative" seems to reoccur as many
    penalties are given more than once.

-   The bar chart indicates that "Labelling Issues" and "Use-by Date"
    cases tend to not repeat.

\newpage

Graphic 2: Stacked Area Plot of Processing Time for Different Types of
Offences Over Time

```{r, echo=FALSE,message=FALSE, warning=FALSE,height=5, width=5}
# Convert to Date format and then calculate the Processing_Time
data <- data %>%
  mutate(Date_Issued = as.Date(Date_Issued, format="%d/%m/%Y"),
         Offence_Date = as.Date(Offence_Date, format="%d/%m/%Y")) %>%
  mutate(Processing_Time = as.numeric(Date_Issued - Offence_Date))
# Check the summary of Processing_Time
#summary(data$Processing_Time)

# Get the 6 most frequent offences
top_offences <- data %>%
  count(Simplified_Nature_of_Offence) %>%
  arrange(desc(n)) %>%
  head(6) %>%
  pull(Simplified_Nature_of_Offence)

# Classify all other offences as 'Others'
data$Offence_Category <- ifelse(data$Simplified_Nature_of_Offence %in% 
                                  top_offences, 
                                data$Simplified_Nature_of_Offence, "Others")

# Plot the box plots for the top 6 offences and 'Others'
ggplot(data, aes(x = Offence_Category, y = Processing_Time, 
                 fill = Offence_Category)) +
  geom_boxplot() +
  scale_fill_brewer(palette = "Set3") +
  theme_minimal() +
  coord_flip() +
  labs(title = "Box Plot of Processing Times for Offences Categories",
       subtitle = "Processing time is the absolute difference in days
       between Date Issued and Offence Date",
       x = "Offence Categories",
       y = "Processing Time (days)") +
  theme(
    plot.title = element_text(hjust = 0.2, face = "bold", size = 12),
    plot.subtitle = element_text(size = 8, hjust = 0.2),  
    axis.text.y = element_text(size = 8), 
    legend.position = "none"
  )

```

**Interpretation of Processing Times across Offence Categories**

-   Processing time is the number of days between the date of offence
    and the date penalty notice issued.

-   Most of the time, for all types of offences, processing times are
    shorter rather than longer.

-   However, in every offence category, there are several cases that
    take much longer time to process, these are the "outliers."

-   Thus, the all the processing time are right-skewed.

-   Other combined categories has a wide range of processing times and a
    lot of variation. This could mean some specific cases take longer to
    process than others on the plot.

\newpage
Graphic 3:Failures in Hand Washing in Top 10 LGAs by Penalty Amount

```{r, echo=FALSE,message=FALSE, warning=FALSE,height=5, width=5}
# Define the keywords that need to be transformed into binary columns
keywords <- c("soap", "single use towels", "warm running water", 
              "easily accessible", 
              "only used for the washing of hands, arms and face",
              "accessible hand washing","easily accessed by food handlers", 
              "previous warning given",  "have hand washing facilities",
              "maintain easily accessible hand washing")

# Filter the data where Nature_of_Offence_Full contains "hand wash"
hand_wash <- data %>%
  filter(str_detect(Nature_of_Offence_Full, regex("hand wash",
                                                  ignore_case = TRUE))) 

# Create new columns for each keyword in the list
for (keyword in keywords) {
  hand_wash <- hand_wash %>%
    mutate(!!keyword := ifelse(str_detect(Nature_of_Offence_Full,
                                   regex(keyword, ignore_case = TRUE)), 1, 0))
}
# Mapping the original keywords to shorter ones
name_mapping <- c(
  "soap" = "soap",
  "single use towels" = "single_use_towels",
  "warm running water" = "warm_water",
  "easily accessible" = "easily_accessible",
  "only used for the washing of hands, arms and face" = "specific_use_only",
  "accessible hand washing" = "accessible_hand_wash",
  "easily accessed by food handlers" = "easy_access_food_handlers",
  "previous warning given" = "previous_warning",
  "have hand washing facilities" = "hand_wash_facilities",
  "maintain easily accessible hand washing" = "maintain_hand_wash_access"
)
# Rename the columns in the hand_wash dataframe
hand_wash <- hand_wash %>%
  rename_with(~ name_mapping[.], all_of(names(name_mapping)))

# binary_columns should have the shorter names, the keys of the name_mapping 
binary_columns <- c(
  "soap", "single_use_towels", "warm_water", "easily_accessible", 
  "specific_use_only", "accessible_hand_wash", "easy_access_food_handlers", 
  "previous_warning", "hand_wash_facilities", "maintain_hand_wash_access"
)
# Aggregate data for LGAs
aggregated_data <- hand_wash %>%
  group_by(Offence_LGA
) %>%
  summarize(
    total_penalty = sum(Amount_Payable, na.rm = TRUE),
    across(all_of(binary_columns), sum, .names = "HW_{col}")
  ) %>%
  arrange(desc(total_penalty)) %>%
  head(10)

melted_data <- aggregated_data %>%
  pivot_longer(
    cols = starts_with("HW_"),
    names_to = "variable",
    values_to = "value"
  )

# Set up the plot
ggplot(melted_data, aes(x = reorder(Offence_LGA, -total_penalty), 
                        y = variable, size = value)) +
  geom_jitter(alpha = 1, width = 0.1, height = 0.1, 
              aes(color = variable)) + 
  scale_size_continuous(range = c(1,9)) +
  scale_color_brewer(palette = "Set3") +  
  theme_light() +  
  labs(
    title = "Failures in Hand Washing in Top 10 LGAs by Penalty Amount", 
    x = "Local Government Area",
    y = "Failure in Hand Washing Facilities",
    caption = "Size of the bubble represents the sum of binary variables"
  ) +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1),
    axis.text.y = element_text(angle = 30, hjust = 1),
    plot.title = element_text(size = 12, face = "bold", hjust = 0.5), 
  ) +
  guides(color=FALSE)
```

**Interpretation:**

-   **Canterbury-Bankstown:** This LGA likely stands out with several
    bubbles of various sizes, depicting numerous instances of
    non-compliance across different aspects of hand washing facilities.

-   **Burwood and Liverpool:** these LGAs would also showcase several
    bubbles. Restaurants in Liverpool may not have hot water in their
    hand washing facilities.

-   **Camden, Blue Mountains and Waverley:** These LGAs would probably
    have the smallest and least amount of bubbles, reflecting the lowest
    total penalties due to fewer non-compliance occurrences in terms of
    hand washing.

-   **Variety in Non-Compliance:** Each LGA might have varying levels
    and types of non-compliance, with some being more prone to specific
    hand washing facility failures than others like Liverpool stands out
    with hot water.

-   **Severity and Frequency of Non-Compliance:** LGAs like
    Canterbury-Bankstown, having more and larger bubbles, depict more
    frequent and severe non-compliance with hand washing facilities.
    

**Important Note:**

The code chunks are hidden as advised. The codes can be found in the R Markdown 
file attached.

**AI Use Acknowledgement**

I have used ChatGPT to revise my writing and debugging codes. It helped make 
the report more professional and concise.

